Transactions Fraud Detection using Machine Learning and Nature Inspired Algorithms

### 2020-05-30 paper/main.tex
Add related works and extend introduction

Data of performed transactions via credit cards are naturally collected by companies and banks to produce statistics or investigate frauds. Therefore, transactions fraud detection can be interpreted as data mining problem, concretely binary classification.

@@

% Required: Úvod / Introduction - aká úloha sa ide riešiť, prečo je dobré ju riešiť

% TODO: co robime, tj, nacrtnut problem atd
% TODO: spomenut, datasety s velkym mnozstvom features, je potrebna feature selection
% TODO: spomenut, ze robi sa to takto takto, ale su aj nejako prirodou inspirovane algoritmy, ktore maju taketo vyhody
% TODO: preco? napr. mzoe chybat domenovat znalost (features) kvoli anonymizacii - nas pripad

@@

@@

>>>>

Data of performed transactions via credit cards are naturally collected by companies and banks to produce statistics or investigate frauds. Therefore, transactions fraud detection can be interpreted as data mining problem, concretely binary classification. Majority of existing datasets of transactions frauds contain a lot of features, however most of them are usually anonymized. Thus, model will rely also on proper feature selection. In such cases, where feature selection is very important part of the problem, powerful metaheuristic search methods like nature inspired algorithms can be used.

@@

@@

Number of transaction frauds is arising alongside with usage of credit cards. To help prevent transaction frauds, common data-mining techniques are available. However, data are in most cases very specific and there is no method suitable to all problems. Addressing that, comparative study of various machine learning techniques, mostly neural networks, for credit card frauds detection was performed~\cite{Sadgali2019}. Common algorithms like Support Vector Machines~\cite{Wiese2009} or Random Forest~\cite{Xuan2018} were also experimented. Another and very specific way to detect frauds is using Genetic programming~\cite{Assis2013}. Genetic programming has proven to be very effective for this problem, achieving significant improvements (up to almost 17\% compared to corporation baseline).

Each transaction can be viewed separately as one unit, but also whole transactions history can be considered. To create model that is resistant to changes in transactions, modeling time series using Long-Short Term Memory (LSTM) recurrent neural networks seems to be the right way~\cite{Wiese2009}.

Publicly available datasets for transactions fraud detection usually contain a lof of features. It is necessary to put special effort on feature selection phase in such cases. Nature inspired algorithms proved to be very powerful for feature selection problems in previous research. When using metaheuristic approaches like nature inspired algorithms, feature selection problem can be interpreted as combinatorial optimization problem~\cite{Wang2016}.

Rough sets are popular when using nature inspired algorithms for feature selection (more about rough sets theory used in data mining in~\cite{Slimani2013}). Rough sets were used in combination with Firefly Algorithm~\cite{Banati2011}, Particle Swarm Optimization~\cite{Wang2007} or Bat Algorithm~\cite{Emary2014}. Majority of works used multiple datasets for their experiments to prove nature inspired algorithms power.

There are also another surveys focused mainly on usage of nature inspired algorithms for feature selection~\cite{Kauser2018,Wang2016,Yang2020}. Further algorithms were researched researched, like Artificial Bees Colony, Cuckoo Search, Ant Colony Optimization or Genetic Algorithms.

@@

% 440 features at the end
